### 2022-10-16 preprint.tex
Deploying to gh-pages from @ gottacatchenall/ms_false_negatives@e411d6fd2bc3a08964824a43ec81f379ae7c2be2 🚀

Here, we assess the effect of false negatives on our ability to make
predictions about interactions, as well as their effect on network
structure. The prevalence of false-negatives in our data is the catalyst
for interaction prediction in the first place, and as a result methods
have been proposed to counteract this bias Poisot \emph{et al.} (2022).
However, it is feasible that the FNR is so high that it could induce too
much noise for an interaction prediction model to detect the signal of
possible interaction between species.

To test this we use the dataset from Hadfield \emph{et al.} (2014) that
describes host-parasite interaction networks sampled across 51 sites,
and the same method as Strydom \emph{et al.} (2021) to extract latent
features for each species in this dataset based on co-occurrence. We
then predict a metaweb (equivalent to predicting true or false for an
interaction for each species pair, effectively a binary classification
problem) from these species-level features using four candidate models
for binary classification---three often used machine-learning (ML)
methods (Boosted Regression Tree (BRT), Random Forest (RF), Decision
Tree (DT)), and one naive model from classic statistics (Logistic
Regression (LR)). Each of the ML models are bootstrap aggregated (or
bagged) with 100 replicates each. We partition the data into 80-20
training-test split, and then seed the training data with false
negatives at varying rates, but crucially do nothing to the test data.
We fit all of these models using MLJ.jl, a high-level Julia framework
for a wide-variety of ML models (Blaom \emph{et al.} 2020). We evaluate
the efficacy of these models using two common measures of binary
classifier performance: the area under the receiver-operator curve
(ROC-AUC) and the area under the precision-recall curve (PR-AUC), for
more details see (Poisot 2022) Here, PR-AUC is slightly more relevant as
it is a better indicator of prediction of false-negatives. The results
of these simulations are shown in fig.~\ref{fig:addedfnr}(A) and (B).

\begin{figure}
\hypertarget{fig:addedfnr}{%
\centering
\includegraphics{./figures/fig3.png}
\caption{(A) The area-under the receiver-operator curve (ROC-AUC) and
(B) The area-under the precision-recall curve (PR-AUC; right) for each
different predictive model (colors/shapes) across a spectrum of the
proportion of added false negatives (x-axis). (C) The mean trophic-level
of all species in a network generated with the niche model across
different species richnesses (colors). For each value of the FNR, the
mean trophic level was computed across 50 replicates. The shaded region
for each line is one standard-deviation across those
replicates.}\label{fig:addedfnr}
}
\end{figure}

One interesting result seen in fig.~\ref{fig:addedfnr}(A) and (B) is
that the ROC-AUC value does not approach random in the same way the
PR-AUC curve does as we increase the added FNR. The reason for this is
that ROC-AUC is fundamentally not as useful a metric in assessing
predictive capacity as PR-AUC. As we keep adding more false-negatives,
the network eventually becomes a zeros matrix, and these models can
still learn to predict ``no-interaction'' for all possible species
pairs, which does far better than random guessing (ROC-AUC
\textgreater{} 0.5) in terms of the true and false positive rates (the
components of ROC-AUC). This highlights a more broad issue of label
class imbalance, meaning there are far more non-interactions than
interactions in data. A full treatment of the importance of
class-balance is outside the scope of this paper, but is explored
in-depth in (Poisot 2022).

Although these ML models are surprisingly performant at link prediction
given their simplicity, there have been several major developments in
the field applying deep-learning methods to many tasks in network
inference and prediction---namely graph-representation learning (GRL,
Khoshraftar \& An (2022)) and graph convolutional networks (Zhang
\emph{et al.} 2019). At this time, these advances can not yet be applied
to ecological networks because they require far more data than we
currently have. We already have lots of features that could be used as
inputs into these models (i.e.~species level data about occurrence,
genomes, abundance, etc.), but our network datasets barely get into the
hundreds of local networks sampled across space and time
tbl.~\ref{tbl:id}. Once we start to get into the thousands, these models
will become more useful, but this can only be done with systematic
monitoring of interactions. This again highlights the need to optimize
our sampling effort to maximize the amount of information contained in
our data given the expensive nature of sampling interactions.

We also consider how the FNR affects network properties. In
fig.~\ref{fig:addedfnr}(C) we see the mean trophic level across networks
simulated using the niche model (as above), across a spectrum of FNR
values. In addition to the clear dependence on richness, we see that
mean trophic level, despite varying widely between niche model
simulations, tends to be relatively robust to false negatives and does
not deviate widely from the true value until very large FNRs,
i.e.~\(p_{fn} > 0.7\).

@@

Species interactions enable the persistence and functioning of
ecosystems, but our understanding of interactions is limited due to the
intrinsic difficulty of sampling. Here we have provided a null model for
the expected number of false-negatives in an interaction dataset. We
demonstrated that we expect many false-negatives in species interaction
datasets purely due to the intrinsic variation of abundances within a
community. We also, for the first time to our knowledge, measured the
strength of association between co-occurrence and interactions (Cazelles
\emph{et al.} 2016) across many empirical systems for the first time,
and found that these positive associations are both very common, and
showed algebraically that they increase the realized FNR. We have also
shown that false-negatives could further impact our ability to both
predict interactions and infer properties of the networks, which
highlights the need for further research into methods for correcting
this bias in existing data.

A better understanding of how false-negatives impact our inference of
network structure and dynamics is prediction of ecological networks by
both using information about individual species interactions and the
structure of metawebs on large scales is a practical necessity.
False-negatives could pose a problem for many forms of inference in
network ecology. For example, inferring the structural or dynamic
stability of a network could be prone to error if the observed network
is not sampled ``enough.'' What exactly ``enough'' means is then
specific to the application, and should be assessed via methods like
those here when designing samples. Further, predictions about network
rewiring (Thompson \& Gonzalez 2017) due to range shifts in response to
climate change could be error-prone without accounting for interactions
that have not been observed but that still may become climatically
infeasible. As is evident from Figure 1(A), we can never guarantee there
are no false-negatives in data. In recent years, there has been interest
toward explicitly accounting for false-negatives in models (Stock
\emph{et al.} 2017; Young \emph{et al.} 2021), and a predictive approach
to networks---rather than expecting our samples to fully capture all
interactions (Strydom \emph{et al.} 2021). As a result, better models
for predicting interactions are needed for interaction networks. This
includes explicitly accounting for observation error (Johnson \&
Larremore 2021)---certain classes of models have been used to reflect
hidden states which account for detection error in occupancy modeling
(Joseph 2020), and could be integrated in the predictive models of
interactions in the future.

A brief caveat here is that we do not consider the rate of
false-positives---in large part false-positives can be explained by
misidentification of species, although this could be a relevant
consideration in some cases. The same logic that we apply to
false-negatives could easily be applied to false-positives, e.g.~that we
can be much more confident that an interaction is a true positive if we
have observed it 50 times rather than only once, and we could similarly
model this using the geometric distribution as in
fig.~\ref{fig:geometric}(A). However, because ecological networks are so
sparse, there are far more negatives than positives in the dataset, and
therefore likely to be far more false-negatives than false-positives in
absolute terms.

@@

>>>>

Here, we assess the effect of false negatives on our ability to make
predictions about interactions, as well as their effect on network
structure. The prevalence of false-negatives in data is the catalyst for
interaction prediction in the first place, and as a result methods have
been proposed to counteract this bias (Stock \emph{et al.} 2017; Poisot
\emph{et al.} 2022). However, it is feasible that the FNR in a given
dataset is so high that it could induce too much noise for an
interaction prediction model to detect the signal of possible
interaction between species.

To test this we use the dataset from Hadfield \emph{et al.} (2014) that
describes host-parasite interaction networks sampled across 51 sites,
and the same method as Strydom \emph{et al.} (2021) to extract latent
features for each species in this dataset based on applying PCA to the
co-occurrence matrix. We then predict a metaweb (equivalent to
predicting true or false for an interaction between each species pair,
effectively a binary classification problem) from these species-level
features using four candidate models for binary classification---three
often used machine-learning (ML) methods (Boosted Regression Tree (BRT),
Random Forest (RF), Decision Tree (DT)), and one naive model from
classic statistics (Logistic Regression (LR)). Each of the ML models are
bootstrap aggregated (or bagged) with 100 replicates each. We partition
the data into 80-20 training-test split, and then seed the training data
with false negatives at varying rates, but crucially do nothing to the
test data. We fit all of these models using MLJ.jl, a high-level Julia
framework for a wide-variety of ML models (Blaom \emph{et al.} 2020). We
evaluate the efficacy of these models using two common measures of
binary classifier performance: the area under the receiver-operator
curve (ROC-AUC) and the area under the precision-recall curve (PR-AUC),
for more details see Poisot (2022). Here, PR-AUC is slightly more
relevant as it is a better indicator of prediction of false-negatives.
The results of these simulations are shown in
fig.~\ref{fig:addedfnr}(A\&B).

\begin{figure}
\hypertarget{fig:addedfnr}{%
\centering
\includegraphics{./figures/fig3.png}
\caption{\textbf{(A)} The area-under the receiver-operator curve
(ROC-AUC) and \textbf{(B)} The area-under the precision-recall curve
(PR-AUC; right) for each different predictive model (colors/shapes)
across a spectrum of the proportion of added false negatives (x-axis).
\textbf{(C)} The mean trophic-level of all species in a network
generated with the niche model across different species richnesses
(colors). For each value of the FNR, the mean trophic level was computed
across 50 replicates. The shaded region for each line is one
standard-deviation across those replicates.}\label{fig:addedfnr}
}
\end{figure}

One interesting result seen in fig.~\ref{fig:addedfnr}(A\&B) is that the
ROC-AUC value does not approach random in the same way the PR-AUC curve
does as we increase the added FNR. The reason for this is that ROC-AUC
is fundamentally not as useful a metric in assessing predictive capacity
as PR-AUC. As we keep adding more false-negatives, the network
eventually becomes a zeros matrix, and these models can still learn to
predict ``no-interaction'' for all possible species pairs, which does
far better than random guessing (ROC-AUC = 0.5) in terms the false
positive rate (one of the components of ROC-AUC). This highlights a more
broad issue of label class imbalance, meaning there are far more
non-interactions than interactions in data. A full treatment of the
importance of class-balance is outside the scope of this paper, but is
explored in-depth in Poisot (2022).

Although these ML models are surprisingly performant at link prediction
given their simplicity, there have been several major developments in
applying deep-learning methods to many tasks in network inference and
prediction---namely graph-representation learning (GRL, Khoshraftar \&
An (2022)) and graph convolutional networks (Zhang \emph{et al.} 2019).
At this time, these advances can not yet be applied to ecological
networks because they require far more data than we currently have. We
already have lots of features that could be used as inputs into these
models (i.e.~species level data about occurrence, genomes, abundance,
etc.), but our network datasets barely get into the hundreds of local
networks sampled across space and time (tbl.~\ref{tbl:id}). Once we
start to get into the thousands, these models will become more useful,
but this can only be done with systematic monitoring of interactions.
This again highlights the need to optimize our sampling effort to
maximize the amount of information contained in our data given the
expensive nature of sampling interactions.

We also consider how the FNR affects network properties. In
fig.~\ref{fig:addedfnr}(C) we see the mean trophic level across networks
simulated using the niche model (as above), across a spectrum of FNR
values. In addition to the clear dependence on richness, we see that
mean trophic level, despite varying widely between niche model
simulations, tends to be relatively robust to false negatives and does
not deviate widely from the true value until very large FNRs,
i.e.~\(p_{fn} > 0.7\). This is not entirely unsurprising. Removing links
randomly from a food-web is effectively the inverse problem of the
emergence of a giant component (more than half of the nodes are in a
connected network) in random graphs (see Li \emph{et al.} (2021) for a
thorough review). The primary difference being that we are removing
edges, not adding them, and thus we are witnessing the dissolution of a
giant component, rather than the emergence of one. Further applications
of percolation theory to the topology of ecological networks could
improve our understanding of how false-negatives impact the inferences
about the structure and dynamics on these networks.

@@

Species interactions enable the persistence and functioning of
ecosystems, but our understanding of interactions is limited due to the
intrinsic difficulty of sampling. Here we have provided a null model for
the expected number of false-negatives in an interaction dataset. We
demonstrated that we expect many false-negatives in species interaction
datasets purely due to the intrinsic variation of abundances within a
community. We also, for the first time to our knowledge, measured the
strength of association between co-occurrence and interactions (Cazelles
\emph{et al.} 2016) across many empirical systems, and found that these
positive associations are both very common, and showed algebraically
that they increase the realized FNR. We have also shown that
false-negatives could further impact our ability to both predict
interactions and infer properties of the networks, which highlights the
need for further research into methods for correcting this bias in
existing data.

A better understanding of how false-negatives impact species interaction
data is a practical necessity---both for inference of network structure
and dynamics, but also for prediction of interactions by using species
level information. False-negatives could pose a problem for many forms
of inference in network ecology. For example, inferring the dynamic
stability of a network could be prone to error if the observed network
is not sampled ``enough.'' What exactly ``enough'' means is then
specific to the application, and should be assessed via methods like
those here when designing samples. Further, predictions about network
rewiring (Thompson \& Gonzalez 2017) due to range shifts in response to
climate change could be error-prone without accounting for interactions
that have not been observed but that still may become climatically
infeasible. As is evident from fig.~\ref{fig:geometric}(A), we can never
guarantee there are no false-negatives in data. In recent years, there
has been interest toward explicitly accounting for false-negatives in
models (Stock \emph{et al.} 2017; Young \emph{et al.} 2021), and a
predictive approach to networks---rather than expecting our samples to
fully capture all interactions (Strydom \emph{et al.} 2021). As a
result, better models for predicting interactions are needed for
interaction networks. This includes explicitly accounting for
observation error (Johnson \& Larremore 2021)---certain classes of
models have been used to reflect hidden states which account for
detection error in occupancy modeling (Joseph 2020), and could be
integrated in the predictive models of interactions in the future.

A brief caveat here is that we do not consider false-positives---in
large part false-positives can be explained by misidentification of
species, although this could be a relevant consideration in some cases.
The same logic that we apply to false-negatives could easily be applied
to false-positives, e.g.~that we can be much more confident that an
interaction is a true positive if we have observed it 50 times rather
than only once, and we could similarly model this using the geometric
distribution as in fig.~\ref{fig:geometric}(A). However, because
ecological networks are so sparse, there are far more negatives than
positives in the dataset, and therefore likely to be far more
false-negatives than false-positives in absolute terms.

@@

\leavevmode\hypertarget{ref-Li2021PerCom}{}%
Li, M., Liu, R.-R., Lü, L., Hu, M.-B., Xu, S. \& Zhang, Y.-C. (2021).
Percolation on complex networks: Theory and application. \emph{Physics
Reports}, Percolation on complex networks: Theory and application, 907,
1--68.
